---
title: "a05"
format: html
---

## Introduction 
This is the Socioecological Monitoring of the Sacramento-San Joaquin Delta. The data collected is at different locations along the delta and types of visitors at said locations. The analysis compiles this data together for ease of interpretation via visualizing the data. 



```{r, warning = FALSE, message=FALSE}
library(readr)
library(dplyr)
library(tidyr)
library(forcats) # makes working with factors easier
library(ggplot2)
library(leaflet) # interactive maps
library(DT) # interactive tables
library(scales) # scale functions for visualization
library(janitor) # expedite cleaning and exploring data
library(viridis) # colorblind friendly color palette
```

```{r}
# read in data from the data directory after manually downloading data 
delta_visits_raw <- read_csv(here::here("data/Socioecological_monitoring_data.csv"))

## Check out column names

colnames(delta_visits_raw)

## Peak at each column and class
glimpse(delta_visits_raw)

## From when to when
range(delta_visits_raw$Date)

## Which time of day?
unique(delta_visits_raw$Time_of_Day)
```

```{r}
delta_visits <- delta_visits_raw %>% 
    janitor::clean_names()
colnames(delta_visits)
```

```{r}
visits_long <- delta_visits %>% 
    pivot_longer(cols = c(sm_boat, med_boat, lrg_boat, bank_angler, scientist, cars),
                 names_to = "visitor_type",
                 values_to = "quantity") %>%
    rename(restore_loc = eco_restore_approximate_location) %>% 
    select(-notes)

## Checking the outcome
head(visits_long)
```

```{r}
daily_visits_loc <- visits_long %>%
    group_by(restore_loc, date, visitor_type) %>% 
    summarise(daily_visits = sum(quantity))
    
head(daily_visits_loc)
```

```{r}
daily_visits_totals <- daily_visits_loc %>% 
    group_by(restore_loc) %>%
    mutate(n = sum(daily_visits)) %>% 
    ungroup() %>%
    mutate(restore_loc = fct_reorder(restore_loc, n))

head(daily_visits_totals)
```

```{r}
levels(daily_visits_totals$restore_loc) ### not alphabetical any more!
```

```{r}
#making the plot 
facet_plot <- ggplot(data = daily_visits_totals,
#naming the plot facet_plot and assigning the data
       aes(x = visitor_type, y = daily_visits,
#assigning the x and y values from the data 
           fill = visitor_type)) +
#assigning the "fill" value from the data that makes the bars 
    geom_col() +
#making a column graph 
      facet_wrap(~restore_loc,
               scales = "free_y",
               ncol = 5,
               nrow = 2) +
#fixing the scale of the graph 
      scale_fill_viridis_d() +
#assigning colors
      labs(x = "Type of visitor",
         y = "Number of Visits",
         title = "Total Number of Visits to Delta Restoration Areas",
         subtitle = "Sum of all visits during study period") +
#labeling the axes and subtitle 
    theme_bw() +
    theme(legend.position = "bottom",
          axis.ticks.x = element_blank(),
          axis.text.x = element_blank())
#assigning the theme, legend, text and ticks position  
facet_plot
#printing the plot 
```

```{r}
locations <- visits_long %>%
    distinct(restore_loc, .keep_all = TRUE) %>%
    select(restore_loc, latitude, longitude)

head(locations)
```

```{r}
datatable(locations)
```

```{r}
#making the map 
leaflet(locations) %>%
#calling the map leaflet (locations)
    addProviderTiles(providers$Esri.WorldImagery) %>%
#adding locations 
    addProviderTiles(providers$CartoDB.PositronOnlyLabels) %>%
#adding location labels 
    addCircleMarkers(
        lng = ~ longitude,
        lat = ~ latitude,
        popup = ~ restore_loc,
        radius = 5,
#adding the circle markers 
        fillColor = "salmon",
        fillOpacity = 1,
#coloring the circle markers 
        stroke = TRUE,
        weight = 0.5,
        color = "white",
        opacity = 1)
#adding the strokes and coloring them 

```

## Discussion
The graph shows the total number of visits to delta restoration. It separates the data by visitor type and what location it is. The colors correspond to the types of visitors. This shows all of the data in one central graph.
The map shows the locations that the samples were collected from. The red dots correspond to locations that were sampled. The text is for the cities that are nearby. The map is also interactive. You can zoom in and click on the dots for the location names. 